---
title: "Spatial Interpolation Overview"
author: "Arsalan Anwari"
date: "2023-06-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

setwd("C:/Users/arsal/Workspaces/thesis_project_f2")

```


# 1. Data preparation
## 1.1 Load libraries 

```{r load_libs, warning=FALSE}

library(jsonlite) # used for wind speed data of knmi station
library(sf) # used to load maps and manipulate spatial features

# used to plot maps and other figures
library(ggplot2)
library(gridExtra) # to create nxn facets for ggplot

library(tidyverse) # used for piping functionality of data manipulation functions

# Used to fix conflicted packages of dplyr
library(conflicted)
conflict_prefer("filter", "dplyr")
conflict_prefer("lag", "dplyr")

library(gstat) # used for interpolation and validation of IDW and kriging models
library(caret) # used for cross validation of other models

```

## 1.2 Load data

```{r load_data}

# load windspeed measurements
df_knmi_windspeed <- read_json("Data/knmi_winspeed_days_2017.json", simplifyVector = TRUE)
df_knmi_station_coords <- read.csv( "Data/knmi_weather_stations.csv", sep=";")

crs_source <- readRDS("Data/source_map_crs.rds") 
sf_netherlands <- st_read("Data/netherlands_reference_map.gpkg")
ls_best_rmse_weights <- readRDS("Data/weights_with_lowest_rmse_per_opt.rds")

```

## 1.3 Clean windspeed data
### 1.3.1 Rename windspeed dataframe columns

```{r rename_df_windspeed}

df_knmi_windspeed <- df_knmi_windspeed %>% 
  select(station_code, date, FG) %>%
  rename(windspeed = FG)

```

### 1.3.3 Find missing entries and na values

```{r find_missing_stations}
missing_rows_stations <- df_knmi_windspeed %>% 
  group_by(station_code) %>%
  summarise(n = n()) %>%
  filter(n < 365) %>% 
  select(station_code)
  
na_rows_stations <- df_knmi_windspeed %>% 
  filter(is.na(windspeed)) %>%
  select(station_code)
  
missing_knmi_stations <- unique(rbind(missing_rows_stations, na_rows_stations))
  
missing_knmi_stations
```

### 1.3.3 filter out rows with missing entries

```{r filter_df_windspeed}
df_knmi_windspeed <- df_knmi_windspeed %>% 
  filter(!(station_code %in% missing_knmi_stations$station_code))

df_knmi_windspeed

```

## 1.4 Merge location data of `df_knmi_station_coords` with windspeed data of `df_knmi_windspeed` into new variable

```{r merge_knmi_data}

df_knmi_station_coords <- df_knmi_station_coords %>%
  select(STN, LON, LAT, NAME) %>%
  rename(station_code = STN, x=LON, y=LAT, station_name=NAME)

df_knmi_stations <- merge(df_knmi_windspeed, df_knmi_station_coords, by = "station_code")

df_knmi_stations

```

## 1.5 Fix datetime and restructure order of dataframe

```{r clean_knmi_data}

df_knmi_stations <- df_knmi_stations %>% 
  mutate(date=strftime(date, "%d-%m-%Y")) %>%
  mutate(date=as.Date(date, "%d-%m-%Y")) %>%
  select(station_code, station_name, date, x, y, windspeed)

```


## 1.6 Aplly source CRS to reference map and create 5km2 grid

```{r create_reference_grid}

sf_netherlands <- sf_netherlands %>%
  st_geometry() %>%
  st_transform(crs_source)

grid_spacing <- 5000

gs_netherlands_grid <- sf_netherlands %>% 
  st_make_grid(
    square = TRUE, 
    what = "polygons",           
    cellsize = c(grid_spacing, grid_spacing)
  )

sf_netherlands_grid <- gs_netherlands_grid %>% st_sf()

sf_netherlands_grid

```

## 1.7 Remove seatiles and other unneccesary tiles

```{r reduce_reference_grid}

sf_netherlands_grid <- sf_netherlands_grid %>%
  mutate(cell_id = 1:length(lengths(gs_netherlands_grid))) # add cell ID

sf_netherlands_grid$n_intersects <- lengths(
  st_intersects(sf_netherlands_grid, sf_netherlands)
)

sf_netherlands_grid <- sf_netherlands_grid %>%
  filter(n_intersects > 0) %>%
  select(geometry, cell_id)
  
sf_netherlands_grid

```

## 1.8 Show map of grids

```{r show_grid_map}

ggplot() +
  geom_sf(data = sf_netherlands, fill="white", col="black") +
  geom_sf(data = sf_netherlands_grid, alpha=0.1) +
  theme_bw()

```



# 2. Observation and receptor preparation
## 2.1 Take single sample of knmi windspeed timeseries as observations

```{r create_sf_observations}

df_observations <- df_knmi_stations %>% 
  filter(date == as.Date("2017-01-01")) %>% 
  select(!date)

sf_observations <- df_observations %>% 
  st_as_sf(coords = c("x", "y")) %>%
  st_set_crs(4326) %>% # WGS-84
  st_transform(crs_source)
  
sf_observations
df_observations

```

## 2.2 Find grid which intersect with observations. 
```{r observation_receptor_instersect}

sf_receptors <- sf_netherlands_grid
sf_receptors$station_codes <- c(st_intersects(sf_netherlands_grid, sf_observations))

```

### 2.2.1 Create backup grid for plotting

```{r create_netherlands_receptor_grid}

sf_receptors_copy <- sf_netherlands_grid
sf_receptors_copy$station_codes <- c(st_intersects(sf_netherlands_grid, sf_observations))
sf_netherlands_receptor_grid <- sf_receptors_copy %>% filter(lengths(station_codes) == 0)

sf_netherlands_receptor_grid

```

## 2.3 Apply found indexes of intersects with windspeed values.

```{r apply_intersects}

# R has a weird bug where using the mean() function over a flattend list of length 1 used  will take the mean of the entire data frame instead of the selection. 
# Therefor two separate search operations are performed, the results combined # with each other and finally merged with the receptor array.

single_intersects <- sf_receptors %>%
  as.data.frame() %>%
  select(cell_id, station_codes) %>%
  filter(lengths(station_codes) == 1) %>%
  mutate(windspeed = df_observations$windspeed[unlist(station_codes)]) %>%
  select(cell_id, windspeed)

multi_intersects <- sf_receptors %>%
  as.data.frame() %>%
  select(cell_id, station_codes) %>%
  filter(lengths(station_codes) > 1) %>%
  mutate(windspeed = mean(df_observations$windspeed[unlist(station_codes)])) %>%
  select(cell_id, windspeed)

all_intersects <- rbind(single_intersects, multi_intersects)

sf_observations <- merge(sf_receptors, all_intersects, by="cell_id")

```

```{r show_intersection_results}

sf_observations

```

## 2.4 Calculate the centroids of the receptors are use that as the geometry instead of polygons. 

```{r calc_receptor_centroids}

sf_receptors <- sf_receptors %>%
  st_set_geometry(st_centroid(st_geometry(sf_receptors)))

sf_observations <- sf_observations %>%
  st_set_geometry(st_centroid(st_geometry(sf_observations)))

```

## 2.5 Set default value of 0 for receptor variables

```{r set_receptor_variable_default}

sf_receptors <- sf_receptors %>% mutate(windspeed = 0)

```

## 2.6 Remove observation cells from receptor grid

```{r remove_obs_cells_from_receptor_grid}

sf_receptors <- sf_receptors %>%
  filter(lengths(station_codes) == 0)

```

## 2.7 Prepare and save data for other R scripts like dynamic model traning routine

```{r ceate_annual_observations}

sf_observations_anual <- df_knmi_stations %>% 
  st_as_sf(coords = c("x", "y")) %>%
  st_set_crs(4326) %>% # WGS-84
  st_transform(crs_source)

```

```{r split_sf_coords}

sf_receptors <- sf_receptors %>%
  mutate(
    x = st_coordinates(.)[,1],
    y = st_coordinates(.)[,2]
  ) 

sf_observations <- sf_observations %>%
  mutate(
    x = st_coordinates(.)[,1],
    y = st_coordinates(.)[,2]
  )

sf_observations_anual <- sf_observations_anual %>%
  mutate(
    x = st_coordinates(.)[,1],
    y = st_coordinates(.)[,2]
  )

```

```{r save_data_for_dynamic_training}

saveRDS(sf_observations, "Data/knmi_windspeed_observations_training_set.rds")
saveRDS(sf_observations_anual, "Data/knmi_windspeed_complete_set.rds")
saveRDS(sf_receptors, "Data/knmi_windspeed_receptors_training_set.rds")

```


# 3. Spatial intepolation methods

A total of 6 spatial interpolation models are tested with "optimal" weights and no weights added. The optimal weights were calculated using the code in `dyanmic-training-routine.R`. 
- In chapter 5 The RMSE and NRMSE of the weighted and non-weighted models are calculated. 
- Chapter 6 compares the best case and average case models by showing their differences in RMSE and NRMSE and by plotting predicted surfaces on a map. 

The optimal weights are calculated such that they perform well on average across all available days in the year 2017. For some models it is possible that omitting the weights will actual yield a lower RMSE for a **specific** day but generally speaking will yield a higher RMSE on average when taking **all** days into account. 

## 3.1 Show best combination of weights and settings for spatial interpolation of training set

```{r show_best_weights_per_model_variation}

ls_best_rmse_weights$trend_surface
ls_best_rmse_weights$mqrbf
ls_best_rmse_weights$idw
ls_best_rmse_weights$od_kriging
ls_best_rmse_weights$u_kriging

```
 
Based on these results best case models will use the following weights in this manner:
- Linear and hyperbolic trend surface using `weights = (1/windspeed) + case_weight`
- MQ-RBF model with an `alpha_seed = 1.166434` and `smoothing_factor = 90.770461`
- IDW model using `nmax=6` and `idp=1.028791`
- Ordinary kriging using spherical variogram using `vgm(base_range + 36.41014, "Sph", base_psill + 0.3641014)` where `base_range` and `base_psill` where pre-calculated by fitting the variogram using `fit.variogram(...)` and `nmax=6`
- Universal kriging with linear and hyperbolic regression model using `psill=0.4390736, range=43.90736, nmax=6` and `psill=0.5031153, range=50.31153, nmax=6`. 

## 3.2 Train spatial interpolation models
### 3.2.1 Trend surface 

```{r trend_surface_train}

model.trend_surface.linear.best_case <- lm(
  formula = windspeed ~ x + y,
  data = sf_observations,
  weights = (1 / windspeed) + 1.151302
)

model.trend_surface.hyperbolic.best_case <- lm(
  formula = windspeed ~ x + y + I(x^2) + (x * y) + I(y^2) + I(x^3) + (I(x^2) * y) + (x * I(y^2)) + I(y^3),
  data = sf_observations,
  weights = (1 / windspeed) + 1.000508
)

model.trend_surface.linear.avg_case <- lm(
  formula = windspeed ~ x + y,
  data = sf_observations
)

model.trend_surface.hyperbolic.avg_case <- lm(
  formula = windspeed ~ x + y + I(x^2) + (x * y) + I(y^2) + I(x^3) + (I(x^2) * y) + (x * I(y^2)) + I(y^3),
  data = sf_observations
)

summary(model.trend_surface.linear.best_case)
summary(model.trend_surface.hyperbolic.best_case)
summary(model.trend_surface.linear.avg_case)
summary(model.trend_surface.hyperbolic.avg_case)

```

### 3.2.2 MQ-RBF

```{r mqrbf_model_load}

source("mqrbf-model-simple.R")
args(mqrbf) # Used to define model
args(predict.mqrbf) # Used to predict surface
args(train.mqrbf) # Used for SLOO-CV 

```

```{r mqrbf_train}

model.mqrbf.best_case <- mqrbf(
    data = sf_observations, 
    alpha_seed = 1.166434,
    smoothing_factor = 90.770461
)

model.mqrbf.avg_case <- mqrbf(data = sf_observations)

summary(model.mqrbf.best_case$alpha_weights)
cat("\n")
summary(model.mqrbf.avg_case$alpha_weights)

```

### 3.2.3 IDW

```{r idw_train}

model.idw.best_case <- gstat(
  formula = windspeed ~ 1,
  data = sf_observations,
  nmax = 6,
  set = list(idp = 1.028791)
)

model.idw.avg_case <- gstat(
  formula = windspeed ~ 1,
  data = sf_observations,
  nmax = 6,
  set = list(idp = 1.0)
)

model.idw.best_case
cat("\n")
model.idw.avg_case
```

### 3.2.4 Ordinary kriging

```{r od_kriging_vgm_train}

vgm.od_kriging <- variogram(windspeed ~ 1, data = sf_observations)
model.od_kriging.vgm.avg_case <- suppressWarnings(fit.variogram(vgm.od_kriging, vgm("Sph")))

best_psill <- model.od_kriging.vgm.avg_case$psill[1] + 0.3641014
best_range <- model.od_kriging.vgm.avg_case$range[2] + 36.41014

model.od_kriging.vgm.best_case <- suppressWarnings(
  fit.variogram(vgm.od_kriging, vgm(best_psill, model = "Sph", best_range))
)

cat(sprintf("
  avg_psill = %f; best_psill = %f
  avg_range = %f; best_range = %f
", 
  model.od_kriging.vgm.avg_case$psill[1], best_psill, 
  model.od_kriging.vgm.avg_case$range[2], best_range
))

```

```{r od_kriging_train}

model.od_kriging.best_case <- gstat(
  formula = windspeed ~ 1,
  data = sf_observations,
  nmax = 6,
  model = model.od_kriging.vgm.best_case
)

model.od_kriging.avg_case <- gstat(
  formula = windspeed ~ 1,
  data = sf_observations,
  nmax = 6,
  model = model.od_kriging.vgm.avg_case
)

```


### 3.2.5 Universal kriging

```{r u_kriging_vgm_train}

vgm.u_kriging <- variogram(windspeed ~ x + y, data = sf_observations)
model.u_kriging.vgm.avg_case <- suppressWarnings(fit.variogram(vgm.u_kriging, vgm("Sph")))

best_psill.linear <- model.u_kriging.vgm.avg_case$psill[1] + 0.4390736
best_range.linear <- model.u_kriging.vgm.avg_case$range[2] + 43.90736
best_psill.hyperbolic <- model.u_kriging.vgm.avg_case$psill[1] + 0.5031153
best_range.hyperbolic <- model.u_kriging.vgm.avg_case$range[2] + 50.31153

model.u_kriging.vgm.best_case.linear <- suppressWarnings(
  fit.variogram(vgm.od_kriging, vgm(best_psill.linear, model = "Sph", best_range.linear))
)

model.u_kriging.vgm.best_case.hyperbolic <- suppressWarnings(
  fit.variogram(vgm.od_kriging, vgm(best_psill.hyperbolic, model = "Sph", best_range.hyperbolic))
)

cat(sprintf("
  [Linear]
  avg_psill = %f; best_psill = %f
  avg_range = %f; best_range = %f
  
  [Hyperbolic]
  avg_psill = %f; best_psill = %f
  avg_range = %f; best_range = %f
", 
  model.u_kriging.vgm.avg_case$psill[1], best_psill.linear, 
  model.u_kriging.vgm.avg_case$range[2], best_range.linear,
  model.u_kriging.vgm.avg_case$psill[1], best_psill.hyperbolic, 
  model.u_kriging.vgm.avg_case$range[2], best_range.hyperbolic
))

```
- The syntax for a `formula` in the `variogram()` function is unconventional. Here `windspeed ~ x+y` simply indicates "variogram optimized for universal kriging" and `windspeed ~ 1` indicates simple or ordinary kriging.  


```{r u_kriging_train}

model.u_kriging.linear.best_case <- gstat(
  formula = windspeed ~ x + y,
  data = sf_observations,
  nmax = 6,
  model = model.u_kriging.vgm.best_case.linear
)

model.u_kriging.hyperbolic.best_case <- gstat(
  formula = windspeed ~ abs(x + y + (x^2) + (x * y) + (y^2) + (x^3) + ((x^2) * y) + (x * (y^2)) + (y^3)),
  data = sf_observations,
  nmax = 6,
  model = model.u_kriging.vgm.best_case.hyperbolic
)

model.u_kriging.linear.avg_case <- gstat(
  formula = windspeed ~ x + y,
  data = sf_observations,
  nmax = 6,
  model = model.u_kriging.vgm.avg_case
)

model.u_kriging.hyperbolic.avg_case <- gstat(
  formula = windspeed ~ abs(x + y + (x^2) + (x * y) + (y^2) + (x^3) + ((x^2) * y) + (x * (y^2)) + (y^3)),
  data = sf_observations,
  nmax = 6,
  model = model.u_kriging.vgm.avg_case
)

```
- The syntax for the `formula` parameter is unconventional in the `gstat()` function. Exponents do not need a `I(x^n)` clause and `abs()` is used to indicate that a higher order polynomial regression model is used without auxiliary variables. To use auxiliary variables for something like co-kriging you need to remove the `abs(....)` and use `log(windspeed)`.    



# 4. Calculate RMSE and NRMSE for each model
## 4.1 Define helper functions 
### 4.1.1 RMSE and NRMSE functions
```{r calc_nrmse}

calc_rmse <- function(residuals) {
  return(sqrt(sum((residuals)^2)/length(residuals)))
}

calc_nrmse <- function(rmse){
  return( rmse / sd(sf_observations$windspeed) )
}

```

### 4.1.2 Decorator for SLOO-CV crossvalidation function for models using `gstat` library
```{r train_gstat_cv}

train.gstat <- function(model){
  { sink(nullfile()); cv <- gstat.cv(model, nfold = nrow(sf_observations), verbose=FALSE); sink(); }  
  return( list(residuals = cv$residual) )
}

```
- The `gstat.cv()` function has the tendency to output all training epochs in the terminal, making for unnecessary long knitted html files. Using `sink(nullfile())` is equivalent to piping all output of a command to `/dev/null`. This removes the output visually in the generated html file.   

## 4.2 Calculate RMSE and NRMSE for each weighted and non weighted model
### 4.2.1 Trend surface

```{r calc_rmse_trend_surface}

# Calc RMSE using SLOO-CV
results.trend_surface.linear.best_case <- train(
  windspeed ~ x + y,
  method = "lm",
  data = sf_observations,
  trControl = trainControl(method = "LOOCV"),
  weights = (1 / windspeed) + 1.151302
)

results.trend_surface.hyperbolic.best_case <- train(
  windspeed ~ x + y + I(x^2) + (x * y) + I(y^2) + I(x^3) + (I(x^2) * y) + (x * I(y^2)) + I(y^3),
  method = "lm",
  data = sf_observations,
  trControl = trainControl(method = "LOOCV"),
  weights = (1 / windspeed) + 1.000508
)

results.trend_surface.linear.avg_case <- train(
  windspeed ~ x + y,
  method = "lm",
  data = sf_observations,
  trControl = trainControl(method = "LOOCV")
)

results.trend_surface.hyperbolic.avg_case <- train(
  windspeed ~ x + y + I(x^2) + (x * y) + I(y^2) + I(x^3) + (I(x^2) * y) + (x * I(y^2)) + I(y^3),
  method = "lm",
  data = sf_observations,
  trControl = trainControl(method = "LOOCV")
)

# Calc NRMSE from RMSE
results.trend_surface.linear.rmse.best_case <- calc_rmse(resid(results.trend_surface.hyperbolic.best_case))
results.trend_surface.linear.nrmse.best_case <- calc_nrmse(results.trend_surface.linear.rmse.best_case)

results.trend_surface.hyperbolic.rmse.best_case <- calc_rmse(resid(results.trend_surface.hyperbolic.best_case))
results.trend_surface.hyperbolic.nrmse.best_case <- calc_nrmse(results.trend_surface.hyperbolic.rmse.best_case)

results.trend_surface.linear.rmse.avg_case <- calc_rmse(resid(results.trend_surface.linear.avg_case))
results.trend_surface.linear.nrmse.avg_case <- calc_nrmse(results.trend_surface.linear.rmse.avg_case)

results.trend_surface.hyperbolic.rmse.avg_case <- calc_rmse(resid(results.trend_surface.hyperbolic.avg_case))
results.trend_surface.hyperbolic.nrmse.avg_case <- calc_nrmse(results.trend_surface.hyperbolic.rmse.avg_case)

```

### 4.2.2 MQ-RBF


```{r calc_rmse_mqrbf}

# Calc RMSE using SLOO-CV
results.mqrbf.best_case <- train.mqrbf(model.mqrbf.best_case)
results.mqrbf.avg_case <- train.mqrbf(model.mqrbf.avg_case)

# Calc NRMSE from RMSE
results.mqrbf.rmse.best_case <- calc_rmse(results.mqrbf.best_case$residuals)
results.mqrbf.nrmse.best_case <- calc_nrmse(results.mqrbf.rmse.best_case)

results.mqrbf.rmse.avg_case <- calc_rmse(results.mqrbf.avg_case$residuals)
results.mqrbf.nrmse.avg_case <- calc_nrmse(results.mqrbf.rmse.avg_case)

```

### 4.2.3 IDW

```{r calc_rmse_idw}

# Calc RMSE using SLOO-CV
results.idw.best_case <- train.gstat(model.idw.best_case)
results.idw.avg_case <- train.gstat(model.idw.avg_case)

# Calc NRMSE from RMSE
results.idw.rmse.best_case <- calc_rmse(results.idw.best_case$residuals)
results.idw.nrmse.best_case <- calc_nrmse(results.idw.rmse.best_case)

results.idw.rmse.avg_case <- calc_rmse(results.idw.avg_case$residuals)
results.idw.nrmse.avg_case <- calc_nrmse(results.idw.rmse.avg_case)
  
```

### 4.2.4 Ordinary kriging

```{r calc_rmse_od_kriging}

# Calc RMSE using SLOO-CV
results.od_kriging.best_case <- train.gstat(model.od_kriging.best_case)
results.od_kriging.avg_case <- train.gstat(model.od_kriging.avg_case)

# Calc NRMSE from RMSE
results.od_kriging.rmse.best_case <- calc_rmse(results.od_kriging.best_case$residuals)
results.od_kriging.nrmse.best_case <- calc_nrmse(results.od_kriging.rmse.best_case)

results.od_kriging.rmse.avg_case <- calc_rmse(results.od_kriging.avg_case$residuals)
results.od_kriging.nrmse.avg_case <- calc_nrmse(results.od_kriging.rmse.avg_case)
  
```

### 4.2.5 Universal kriging

```{r calc_rmse_od_kriging}

# Calc RMSE using SLOO-CV
results.u_kriging.linear.best_case <- train.gstat(model.u_kriging.linear.best_case)
results.u_kriging.hyperbolic.best_case <- train.gstat(model.u_kriging.hyperbolic.best_case)

results.u_kriging.linear.avg_case <- train.gstat(model.u_kriging.linear.avg_case)
results.u_kriging.hyperbolic.avg_case <- train.gstat(model.u_kriging.hyperbolic.avg_case)

# Calc NRMSE from RMSE
results.u_kriging.linear.rmse.best_case <- calc_rmse(results.u_kriging.linear.best_case$residuals)
results.u_kriging.linear.nrmse.best_case <- calc_nrmse(results.u_kriging.linear.rmse.best_case)

results.u_kriging.hyperbolic.rmse.best_case <- calc_rmse(results.u_kriging.hyperbolic.best_case$residuals)
results.u_kriging.hyperbolic.nrmse.best_case <- calc_nrmse(results.u_kriging.hyperbolic.rmse.best_case)

results.u_kriging.linear.rmse.avg_case <- calc_rmse(results.u_kriging.linear.avg_case$residuals)
results.u_kriging.linear.nrmse.avg_case <- calc_nrmse(results.u_kriging.linear.rmse.avg_case)

results.u_kriging.hyperbolic.rmse.avg_case <- calc_rmse(results.u_kriging.hyperbolic.avg_case$residuals)
results.u_kriging.hyperbolic.nrmse.avg_case <- calc_nrmse(results.u_kriging.hyperbolic.rmse.avg_case)

  
```

## 4.3 Combine and show RMSE and NRMSE results of models

```{r compare_rmse_nrmse_per_model}

model_names <- c(
  "Trend surface (1st order)",
  "Trend surface (3rd order)",
  "MQ-RBF",
  "IDW",
  "Ordinary kriging",
  "Universal kriging (1st order)",
  "Universal kriging (3rd order)"
)

best_case_weights <- c(
  "case_weight = 1.151302",
  "case_weight = 1.000508",
  "alpha_seed = 1.166434; smoothing_factor = 90.770461",
  "nmax=6; idp=1.028791",
  "vgm=\"Sph\"; psill=348.3421; range=40943.12; nmax=6",
  "vgm=\"Sph\"; psill=188.709; range=13714.15; nmax=6",
  "vgm=\"Sph\"; psill=188.7081; range=13714.12; nmax=6"
)

best_case_rmse_values <- c(
  results.trend_surface.linear.rmse.best_case,
  results.trend_surface.hyperbolic.rmse.best_case,
  results.mqrbf.rmse.best_case,
  results.idw.rmse.best_case,
  results.od_kriging.rmse.best_case,
  results.u_kriging.linear.rmse.best_case,
  results.u_kriging.hyperbolic.rmse.best_case
)

best_case_nrmse_values <- c(
  results.trend_surface.linear.nrmse.best_case,
  results.trend_surface.hyperbolic.nrmse.best_case,
  results.mqrbf.nrmse.best_case,
  results.idw.nrmse.best_case,
  results.od_kriging.nrmse.best_case,
  results.u_kriging.linear.nrmse.best_case,
  results.u_kriging.hyperbolic.nrmse.best_case
)

avg_case_rmse_values <- c(
  results.trend_surface.linear.rmse.avg_case,
  results.trend_surface.hyperbolic.rmse.avg_case,
  results.mqrbf.rmse.avg_case,
  results.idw.rmse.avg_case,
  results.od_kriging.rmse.avg_case,
  results.u_kriging.linear.rmse.avg_case,
  results.u_kriging.hyperbolic.rmse.avg_case
)

avg_case_nrmse_values <- c(
  results.trend_surface.linear.nrmse.avg_case,
  results.trend_surface.hyperbolic.nrmse.avg_case,
  results.mqrbf.nrmse.avg_case,
  results.idw.nrmse.avg_case,
  results.od_kriging.nrmse.avg_case,
  results.u_kriging.linear.nrmse.avg_case,
  results.u_kriging.hyperbolic.nrmse.avg_case
)

results.all_models <- 
  data.frame(
    model_name = model_names,
    best_case_weight = best_case_weights,
    rmse.best_case = best_case_rmse_values,
    rmse.avg_case = avg_case_rmse_values,
    nrmse.best_case = best_case_nrmse_values,
    nrmse.avg_case = avg_case_nrmse_values
  )

```






# 5. Compare models
## 5.1 Show difference in RMSE and NRMSE between best case and average case models

```{r show_rmse_nrmse_results}

results.all_models
results.all_models %>% select(model_name, rmse.best_case, rmse.avg_case, nrmse.best_case, nrmse.avg_case)

```

## 5.2 Generate surfaces
### 5.2.1 Trend surface
```{r trend_surface_predict}

preds.trend_surface.linear.best_case <- predict(model.trend_surface.linear.best_case, sf_receptors)
preds.trend_surface.hyperbolic.best_case <- predict(model.trend_surface.hyperbolic.best_case, sf_receptors)

preds.trend_surface.linear.avg_case <- predict(model.trend_surface.linear.avg_case, sf_receptors)
preds.trend_surface.hyperbolic.avg_case <- predict(model.trend_surface.hyperbolic.avg_case, sf_receptors)

cat("\n preds.trend_surface.linear.best_case \n")
summary(preds.trend_surface.linear.best_case)

cat("\n preds.trend_surface.hyperbolic.best_case \n")
summary(preds.trend_surface.hyperbolic.best_case)

cat("\n preds.trend_surface.linear.avg_case \n")
summary(preds.trend_surface.linear.avg_case)

cat("\n preds.trend_surface.hyperbolic.avg_case \n")
summary(preds.trend_surface.hyperbolic.avg_case)

```

### 5.2.2 MQ-RBF

```{r mqrbf_predict}

preds.mqrbf.best_case <- predict.mqrbf(model.mqrbf.best_case, sf_receptors)
preds.mqrbf.avg_case <- predict.mqrbf(model.mqrbf.avg_case, sf_receptors)

cat("\n preds.mqrbf.best_case \n")
summary(preds.mqrbf.best_case)

cat("\n preds.mqrbf.avg_case \n")
summary(preds.mqrbf.avg_case)

```

### 5.2.3 IDW

```{r idw_predict}

preds.idw.best_case <- predict(model.idw.best_case, sf_receptors)$var1.pred
preds.idw.avg_case <- predict(model.idw.avg_case, sf_receptors)$var1.pred

cat("\n preds.idw.best_case \n")
summary(preds.idw.best_case)

cat("\n preds.idw.avg_case \n")
summary(preds.idw.avg_case)

```

### 5.2.4 Ordinary kriging

```{r od_kriging_predict}

preds.od_kriging.best_case <- predict(model.od_kriging.best_case, sf_receptors)$var1.pred
preds.od_kriging.avg_case <- predict(model.od_kriging.avg_case, sf_receptors)$var1.pred

cat("\n preds.od_kriging.best_case \n")
summary(preds.od_kriging.best_case)

cat("\n preds.od_kriging.avg_case \n")
summary(preds.od_kriging.avg_case)

```

### 5.2.5 Universal kriging

```{r od_kriging_predict}

preds.u_kriging.linear.best_case <- predict(model.u_kriging.linear.best_case, sf_receptors)$var1.pred
preds.u_kriging.hyperbolic.best_case <- predict(model.u_kriging.hyperbolic.best_case, sf_receptors)$var1.pred

preds.u_kriging.linear.avg_case <- predict(model.u_kriging.linear.avg_case, sf_receptors)$var1.pred
preds.u_kriging.hyperbolic.avg_case <- predict(model.u_kriging.hyperbolic.avg_case, sf_receptors)$var1.pred

cat("\n preds.u_kriging.linear.best_case \n")
summary(preds.u_kriging.linear.best_case)

cat("\n preds.u_kriging.hyperbolic.best_case \n")
summary(preds.u_kriging.hyperbolic.best_case)

cat("\n preds.u_kriging.linear.avg_case \n")
summary(preds.u_kriging.linear.avg_case)

cat("\n preds.u_kriging.hyperbolic.avg_case \n")
summary(preds.u_kriging.hyperbolic.avg_case)

```






## 5.3 Show surfaces
### 5.3.1 Define function to generate map plot

```{r create_prediction_map}

create_prediction_map <- function(observation_variables, title_text, subtitle_text, caption_text){

  sf_netherlands_receptor_grid$windspeed <- 0
  sf_netherlands_receptor_grid$windspeed <- as.factor(ceiling(observation_variables / 10))

  map_plot <- ggplot() +
    geom_sf(
      data = sf_netherlands_receptor_grid,
      aes(fill = windspeed)
    ) +
    scale_fill_grey(start = 0.1, end = 0.9) +
    labs(
      title = title_text, 
      subtitle = subtitle_text, 
      fill = "windspeed (m/s)", 
      caption = caption_text
    ) +
    theme_bw() +
    theme(legend.position = "bottom", 
          legend.direction = "horizontal", 
          axis.text.x=element_blank(),
          axis.text.y=element_blank()
    )
  
  return(map_plot)
}

```

### 5.3.2 Generate map plots

```{r create_best_case_map_plots}

map_plot.trend_surface.linear.best_case <- create_prediction_map(
  preds.trend_surface.linear.best_case,
  title_text = "Trend surface (1st order)",
  subtitle_text = "[best_case]",
  caption_text = "case_weight = 1.151302"
)

map_plot.trend_surface.hyperbolic.best_case <- create_prediction_map(
  preds.trend_surface.hyperbolic.best_case,
  title_text = "Trend surface (1st order)",
  subtitle_text = "[best_case]",
  caption_text = "case_weight = 1.000508"
)

map_plot.mqrbf.best_case <- create_prediction_map(
  preds.mqrbf.best_case,
  title_text = "MQ-RBF",
  subtitle_text = "[best_case]",
  caption_text = "alpha_seed = 1.166434; smoothing_factor = 90.770461"
)

map_plot.idw.best_case <- create_prediction_map(
  preds.idw.best_case,
  title_text = "IDW",
  subtitle_text = "[best_case]",
  caption_text = "nmax=6; idp=1.028791"
)

map_plot.od_kriging.best_case <- create_prediction_map(
  preds.od_kriging.best_case,
  title_text = "Ordinary Kriging",
  subtitle_text = "[best_case]",
  caption_text = "vgm=\"Sph\"; psill=348.706182; range=40979.529490; nmax=6"
)

map_plot.u_kriging.linear.best_case <- create_prediction_map(
  preds.u_kriging.linear.best_case,
  title_text = "Universal Kriging (1st order)",
  subtitle_text = "[best_case]",
  caption_text = "vgm=\"Sph\"; psill=292.601287; range=40987.026710; nmax=6"
)

map_plot.u_kriging.hyperbolic.best_case <- create_prediction_map(
  preds.u_kriging.hyperbolic.best_case,
  title_text = "Universal Kriging (3rd order)",
  subtitle_text = "[best_case]",
  caption_text = "vgm=\"Sph\"; psill=292.665329; range=40993.430880; nmax=6"
)

```

```{r create_avg_case_map_plots}

map_plot.trend_surface.linear.avg_case <- create_prediction_map(
  preds.trend_surface.linear.avg_case,
  title_text = "Trend surface (1st order)",
  subtitle_text = "[avg_case]",
  caption_text = "case_weight = 1.0"
)

map_plot.trend_surface.hyperbolic.avg_case <- create_prediction_map(
  preds.trend_surface.hyperbolic.avg_case,
  title_text = "Trend surface (1st order)",
  subtitle_text = "[avg_case]",
  caption_text = "case_weight = 1.0"
)

map_plot.mqrbf.avg_case <- create_prediction_map(
  preds.mqrbf.avg_case,
  title_text = "MQ-RBF",
  subtitle_text = "[avg_case]",
  caption_text = "alpha_seed = 1.0; smoothing_factor = 0.0"
)

map_plot.idw.avg_case <- create_prediction_map(
  preds.idw.avg_case,
  title_text = "IDW",
  subtitle_text = "[avg_case]",
  caption_text = "nmax=6; idp=1.0"
)

map_plot.od_kriging.avg_case <- create_prediction_map(
  preds.od_kriging.avg_case,
  title_text = "Ordinary Kriging",
  subtitle_text = "[avg_case]",
  caption_text = "vgm=\"Sph\"; psill=348.342080; range=40943.119350; nmax=6"
)

map_plot.u_kriging.linear.avg_case <- create_prediction_map(
  preds.u_kriging.linear.avg_case,
  title_text = "Universal Kriging (1st order)",
  subtitle_text = "[avg_case]",
  caption_text = "vgm=\"Sph\"; psill=292.162214; range=40943.119350; nmax=6"
)

map_plot.u_kriging.hyperbolic.avg_case <- create_prediction_map(
  preds.u_kriging.hyperbolic.avg_case,
  title_text = "Universal Kriging (3rd order)",
  subtitle_text = "[avg_case]",
  caption_text = "vgm=\"Sph\"; psill=292.162214; range=40943.119350; nmax=6"
)

```




### 5.3.3 Combine map plots into facets.

- Create separate facets of sizes (2x1) comparing best case with average case surfaces. 

```{r combine_map_plots}

facet.trend_surface.linear <- arrangeGrob(
  map_plot.trend_surface.linear.best_case, map_plot.trend_surface.linear.avg_case,
  ncol = 2, nrow = 1
)

facet.trend_surface.hyperbolic <- arrangeGrob(
  map_plot.trend_surface.hyperbolic.best_case, map_plot.trend_surface.hyperbolic.avg_case,
  ncol = 2, nrow = 1
)

facet.mqrbf <- arrangeGrob(
  map_plot.mqrbf.best_case, map_plot.mqrbf.avg_case,
  ncol = 2, nrow = 1
)

facet.idw <- arrangeGrob(
  map_plot.idw.best_case, map_plot.idw.avg_case,
  ncol = 2, nrow = 1
)

facet.od_kriging <- arrangeGrob(
  map_plot.od_kriging.best_case, map_plot.od_kriging.avg_case,
  ncol = 2, nrow = 1
)

facet.u_kriging.linear <- arrangeGrob(
  map_plot.u_kriging.linear.best_case, map_plot.u_kriging.linear.avg_case,
  ncol = 2, nrow = 1
)

facet.u_kriging.hyperbolic <- arrangeGrob(
  map_plot.u_kriging.hyperbolic.best_case, map_plot.u_kriging.hyperbolic.avg_case,
  ncol = 2, nrow = 1
)


```


### 5.3.4 Show facets with combined map plots

```{r show_facet_trend_surface_linear}

grid.arrange(facet.trend_surface.linear)

```

```{r show_facet_trend_surface_hyperbolic}

grid.arrange(facet.trend_surface.hyperbolic)

```

```{r show_facet_mqrbf}

grid.arrange(facet.mqrbf)

```

```{r show_facet_idw}

grid.arrange(facet.idw)

```

```{r show_facet_od_kriging}

grid.arrange(facet.od_kriging)

```

```{r show_facet_u_kriging_linear}

grid.arrange(facet.u_kriging.linear)

```

```{r show_facet_u_kriging_hyperbolic}

grid.arrange(facet.u_kriging.hyperbolic)

```

```{r show_facet_all_models}

map_plot.trend_surface.linear.best_case
map_plot.trend_surface.hyperbolic.best_case
map_plot.mqrbf.best_case
map_plot.idw.best_case
map_plot.od_kriging.best_case
map_plot.u_kriging.linear.best_case
map_plot.u_kriging.hyperbolic.best_case

```


