---
title: "thesis_code"
author: "Arsalan Anwari"
date: "2023-06-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Thesis code

## 1. Data preparation
### 1.1 Load libraries 

```{r load_libs, echo=FALSE, warning=FALSE}

library(jsonlite) # used for windspeed data of knmi station
library(sf) # used to load maps and manipulate spatial features

library(ggplot2) # used to plot maps and other figures
library(tidyverse) # used for piping functioality of data manupilation functions

# used for interpolation and validation of idw and kriging models
library(gstat) 

# used for interpolation of RBF models
library(fields)
library(Rcpp)
library(RcppArmadillo)

library(caret) # used for cross validation of other models

```

### 1.2 Load data

```{r load_data, echo=FALSE}

# load windspeed measurements
df_knmi_windspeed <- read_json("Data/knmi_winspeed_days_2017.json", simplifyVector = TRUE)
df_knmi_station_coords <- read.csv( "Data/knmi_weather_stations.csv", sep=";")

sf_source_brpg <- readRDS("Data/brpg_geom_centroids.rds") 
sf_netherlands <- st_read("Data/NL_Country_Boundary.gpkg")

```

### 1.3 Clean windspeed data
#### 1.3.1 Rename windspeed dataframe columns

```{r rename_df_windspeed}

df_knmi_windspeed <- df_knmi_windspeed %>% 
  select(station_code, date, FG) %>%
  rename(windspeed = FG)

```

#### 1.3.3 Find missing entries and na values

```{r find_missing_stations}
missing_rows_stations <- df_knmi_windspeed %>% 
  group_by(station_code) %>%
  summarise(n = n()) %>%
  filter(n < 365) %>% 
  select(station_code)
  
na_rows_stations <- df_knmi_windspeed %>% 
  filter(is.na(windspeed)) %>%
  select(station_code)
  
missing_knmi_stations <- unique(rbind(missing_rows_stations, na_rows_stations))
  
missing_knmi_stations
```

#### 1.3.3 filter out rows with missing entries

```{r filter_df_windspeed}
df_knmi_windspeed <- df_knmi_windspeed %>% 
  filter(!(station_code %in% missing_knmi_stations$station_code))

df_knmi_windspeed

```

### 1.4 Merge location data of `df_knmi_station_coords` with windspeed data of `df_knmi_windspeed` into new variable

```{r merge_knmi_data}

df_knmi_station_coords <- df_knmi_station_coords %>%
  select(STN, LON, LAT, NAME) %>%
  rename(station_code = STN, x=LON, y=LAT, station_name=NAME)

df_knmi_stations <- merge(df_knmi_windspeed, df_knmi_station_coords, by = "station_code")

df_knmi_stations

```

### 1.5 Fix datetime and restructure order of dataframe

```{r clean_knmi_data}

df_knmi_stations <- df_knmi_stations %>% 
  mutate(date=strftime(date, "%d-%m-%Y")) %>%
  mutate(date=as.Date(date, "%d-%m-%Y")) %>%
  select(station_code, station_name, date, x, y, windspeed)

```


### 1.6 Take CRS of source map

```{r get_source_map_crs}

crs_source <- st_crs(sf_source_brpg)

```

### 1.7 Aplly source CRS to reference map and create 5km2 grid

```{r create_reference_grid}

sf_netherlands <- sf_netherlands %>%
  st_geometry() %>%
  st_transform(crs_source)

grid_spacing <- 5000

gs_netherlands_grid <- sf_netherlands %>% 
  st_make_grid(
    square = TRUE, 
    what = "polygons",           
    cellsize = c(grid_spacing, grid_spacing)
  )

sf_netherlands_grid <- gs_netherlands_grid %>% st_sf()

sf_netherlands_grid

```

### 1.8 Remove seatiles and other unneccesary tiles

```{r reduce_reference_grid}

sf_netherlands_grid <- sf_netherlands_grid %>%
  mutate(cell_id = 1:length(lengths(gs_netherlands_grid))) # add cell ID

sf_netherlands_grid$n_intersects <- lengths(
  st_intersects(sf_netherlands_grid, sf_netherlands)
)

sf_netherlands_grid <- sf_netherlands_grid %>%
  filter(n_intersects > 0) %>%
  select(geometry, cell_id)
  
sf_netherlands_grid

```

### 1.9 Show map of grids

```{r show_grid_map}

ggplot() +
  geom_sf(data = sf_netherlands, fill="white", col="black") +
  geom_sf(data = sf_netherlands_grid, alpha=0.1) +
  theme_bw()

```





## 2. Observation and receptor preparation
### 2.1 Take single sample of knmi windspeed timeseries as observations

```{r create_sf_observations}

df_observations <- df_knmi_stations %>% 
  filter(date == as.Date("2017-01-01")) %>% 
  select(!date)

sf_observations <- df_observations %>% 
  st_as_sf(coords = c("x", "y")) %>%
  st_set_crs(4326) %>% # WGS-84
  st_transform(crs_source)
  
sf_observations
df_observations

```

### 2.2 Find grid which intersect with observations. 
```{r observation_receptor_instersect}

sf_receptors <- sf_netherlands_grid
sf_receptors$station_codes <- c(st_intersects(sf_netherlands_grid, sf_observations))

```

### 2.2.1 Create backup grid for plotting

```{r create_netherlands_receptor_grid}

sf_receptors_copy <- sf_netherlands_grid
sf_receptors_copy$station_codes <- c(st_intersects(sf_netherlands_grid, sf_observations))
sf_netherlands_receptor_grid <- sf_receptors_copy %>% filter(lengths(station_codes) == 0)

sf_netherlands_receptor_grid

```

### 2.3 Apply found indexes of intersects with windspeed values.

```{r apply_intersects}

# R has a weird bug where using the mean() function over a flattend list of length 1 used  will take the mean of the entire data frame instead of the selection. 
# Therefor two separate search operations are performed, the results combined # with each other and finally merged with the receptor array.

single_intersects <- sf_receptors %>%
  as.data.frame() %>%
  select(cell_id, station_codes) %>%
  filter(lengths(station_codes) == 1) %>%
  mutate(windspeed = df_observations$windspeed[unlist(station_codes)]) %>%
  select(cell_id, windspeed)

multi_intersects <- sf_receptors %>%
  as.data.frame() %>%
  select(cell_id, station_codes) %>%
  filter(lengths(station_codes) > 1) %>%
  mutate(windspeed = mean(df_observations$windspeed[unlist(station_codes)])) %>%
  select(cell_id, windspeed)

all_intersects <- rbind(single_intersects, multi_intersects)

sf_observations <- merge(sf_receptors, all_intersects, by="cell_id")

```

```{r show_intersection_results}

sf_observations

```

### 2.4 Calculate the centroids of the receptors are use that as the geometry instead of polygons. 

```{r calc_receptor_centroids}

sf_receptors <- sf_receptors %>%
  st_set_geometry(st_centroid(st_geometry(sf_receptors)))

sf_observations <- sf_observations %>%
  st_set_geometry(st_centroid(st_geometry(sf_observations)))

```

### 2.5 Set default value of 0 for receptor variables

```{r set_receptor_variable_default}

sf_receptors <- sf_receptors %>% mutate(windspeed = 0)

```

### 2.6 Remove observation cells from receptor grid

```{r remove_obs_cells_from_receptor_grid}

sf_receptors <- sf_receptors %>%
  filter(lengths(station_codes) == 0)

```

## 3 Perform interpolation
### 3.0 Helper functions and data

```{r plot_observations}

show_prediction_map <- function(observation_variables){
  sf_netherlands_receptor_grid$windspeed <- 0
  sf_netherlands_receptor_grid$windspeed <- observation_variables

  map_plot <- ggplot() +
    geom_sf(data = sf_netherlands_receptor_grid, aes(fill=windspeed))
  
  print(map_plot)
}

```


```{r calc_nrmse}

calc_rmse <- function(residuals) {
  return(sqrt(sum((residuals)^2)/length(residuals)))
}

calc_nrmse <- function(rmse, observations){
  max <- max(observations)
  min <- min(observations)
  return (1.0 - ( rmse / (max - min) ))
}

```

```{r split_sf_coords_to_df}

df_receptors <- sf_receptors %>%
  mutate(
    x = st_coordinates(.)[,1],
    y = st_coordinates(.)[,2]
  ) %>% 
  as.data.frame() %>%
  select(cell_id, x, y, !geometry)

df_observations <- sf_observations %>%
  mutate(
    x = st_coordinates(.)[,1],
    y = st_coordinates(.)[,2]
  ) %>% 
  as.data.frame() %>%
  select(cell_id, x, y, !geometry)

```


### 3.1 Trend surface 
#### 3.1.1 Train nth order trend models

```{r trend_models_train}

trend_surface.first_order.model <- lm(
  formula = windspeed ~ x + y,
  data = df_observations
)

trend_surface.third_order.model <- lm(
  formula = windspeed ~ x + y + I(x^2) + (x * y) + I(y^2) + I(x^3) + (I(x^2) * y) + (x * I(y^2)) + I(y^3),
  data = df_observations
)

```

#### 3.1.2 Predict surface with models

```{r trend_models_predict}

trend_surface.first_order.preds <- as.data.frame(predict(trend_surface.first_order.model, df_receptors))
trend_surface.third_order.preds <- as.data.frame(predict(trend_surface.third_order.model, df_receptors))



```

```{r trend_models_rename_predict}

colnames(trend_surface.first_order.preds) <- c("windspeed")
colnames(trend_surface.third_order.preds) <- c("windspeed")

```

#### 3.1.3 Show results in map

```{r show_first_order_trend_map_results}

show_prediction_map(trend_surface.first_order.preds$windspeed)

```
```{r show_third_order_trend_map_results}

show_prediction_map(trend_surface.third_order.preds$windspeed)

```


#### 3.1.4 Validate results

```{r perform_cv_on_trend_surface_models}

trend_surface.first_order.model_cv <- train(
  windspeed ~ x + y,
  method = "lm",
  data = df_observations,
  trControl = trainControl(method = "LOOCV")
)

trend_surface.third_order.model_cv <- train(
  windspeed ~ x + y + I(x^2) + (x * y) + I(y^2) + I(x^3) + (I(x^2) * y) + (x * I(y^2)) + I(y^3),
  method = "lm",
  data = df_observations,
  trControl = trainControl(method = "LOOCV")
)

```

```{r calc_rsme_trend_surface}

trend_surface.first_order.results.rmse <- (trend_surface.first_order.model_cv$results$RMSE)

trend_surface.first_order.results.nrmse <- calc_nrmse(
  trend_surface.first_order.results.rmse, 
  df_observations$windspeed
)

trend_surface.third_order.results.rmse <- (trend_surface.third_order.model_cv$results$RMSE)
trend_surface.third_order.results.nrmse <- calc_nrmse(
  trend_surface.third_order.results.rmse, 
  df_observations$windspeed
)

print(sprintf(
  "First order trend: rmse=%f nrmse=%f", 
  trend_surface.first_order.results.rmse, 
  trend_surface.first_order.results.nrmse
))

print(sprintf(
  "Third order trend: rmse=%f nrmse=%f", 
  trend_surface.third_order.results.rmse, 
  trend_surface.third_order.results.nrmse
))

```

### 3.2 Radial Basis function
#### 3.2.1 Train MQ-RBF weights based on observations
```{r define_mqrbf_weight_function}

# Calculate RBF weights
calc_mqrbf_alpha_weights <- function(x, y, z, eps) {
  n <- length(x)
  K <- matrix(0, n, n)
  for (i in 1:n) {
    for (j in 1:n) {
      K[i,j] <- sqrt(1 + eps * ((sqrt((x[i]-x[j])^2 + (y[i]-y[j])^2))^2) ) 
    }
  }
  alpha <- solve(K) %*% z
  return(as.vector(alpha))
}

```

```{r calc_mqrbf_weights}

mqrbf.weights <- calc_mqrbf_alpha_weights(
    df_observations$x, df_observations$y, 
    df_observations$windspeed, 1.543523
)

head(mqrbf.weights)
```


#### 3.2.2 Predict surface with MQ-RBF model 

```{r define_mqrbf_model}

calc_mqrbf_distance <- function(r, c){
  return ( sqrt( (r^2) + (c^2) ) )
}

calc_mqrbf_surface <- function(
  x_receptors, y_receptors, 
  x_observations, y_observations,
  alpha_weights, smoothing_factor
){
  
  n_i <- length(x_receptors)
  n_j <- length(x_observations)
  
  # value of 0.05 is to ensure that a float is used and not int. 
  # Values ({0.0 ... 0.009} == 0) in rep() function of R.
  z <- rep(0.01, n_i)  
  
  for(i in 1:n_i) {
    x_new <- x_receptors[i]
    y_new <- y_receptors[i]
    z_new <- 0
    for(j in 1:n_j) {
      weight_j <- alpha_weights[j]
  
      x_j <- x_observations[j]
      y_j <- y_observations[j]
  
      r <- sqrt( ((x_j-x_new)^2) + ((y_j-y_new)^2) )
      m_distance <- calc_mqrbf_distance(r, smoothing_factor)
  
      z_new <- z_new + (weight_j * m_distance)
    }
    z[i] <- z_new
  }
  
  return(z)
  
}

```

```{r predict_mqrbf_surface}

mqrbf.preds <- calc_mqrbf_surface(
  df_receptors$x, df_receptors$y,
  df_observations$x, df_observations$y,
  mqrbf.weights, 0.234334
)

```

#### 3.2.3 Show results in map

```{r show_mqrbf_surface}

show_prediction_map(mqrbf.preds)

```



#### 3.2.4 Validate results

```{r define_mqrbf_cross_validation_method}

calc_mqrbf_loocv_residuals <- function(
    x_observations, y_observations, z_observations,
    training_weight, smoothing_factor
){
  n <- length(x_observations)
  residuals <- rep(0.01, n)
  
  for(i in 1:n) {
    x_train <- x_observations[-i]
    y_train <- y_observations[-i]
    z_train <- z_observations[-i]
    
    x_new <- as.vector(x_observations[i])
    y_new <- as.vector(y_observations[i])
    
    alpha_weights <- calc_mqrbf_alpha_weights(x_train, y_train, z_train, training_weight)
    
    z_new <- calc_mqrbf_surface(
      x_new, y_new,
      x_train, y_train,
      alpha_weights, smoothing_factor
    )
    
    residuals[i] <- (z_observations[i] - z_new)
  }
  
  return(residuals)
}

```

```{r calc_mqrbf_nrmse}

mqrbf.cv.residuals <- calc_mqrbf_loocv_residuals(
  df_observations$x, df_observations$y, df_observations$windspeed,
  1.543523, 0.234334
)

mqrbf.cv.rsme <- calc_rmse(mqrbf.cv.residuals)
mqrbf.cv.nrsme <- calc_nrmse(mqrbf.cv.rsme, df_observations$windspeed)

```

```{r show_mqrbf_nrmse}

print(sprintf(
  "MQ-RBF [eps=1.543523, c=0.234334]: rmse=%f nrmse=%f", 
  mqrbf.cv.rsme, 
  mqrbf.cv.nrsme
))

```
### 3.3 Inverse Distance weighted
#### 3.3.1 Train idw model with `n_neighbors=5` and `idp=1.5`.

```{r train_idw_model}

idw_model.model <- gstat(
  formula = windspeed ~ 1,
  data = sf_observations,
  nmax = 5,
  set = list(idp = 1.5)
)

```

#### 3.3.2 Predict surface with idw model

```{r pred_idw_model}

idw_model.preds <- as.data.frame(predict(idw_model.model, sf_receptors))

```

#### 3.3.3 Show results in map

```{r results_idw_model}

show_prediction_map(idw_model.preds$var1.pred)

```

#### 3.3.4 Validate results

```{r validate_idw_model, echo=FALSE, results='hide', message=FALSE, warning=FALSE}

# if nfolds = nrow(observations) than gstat.cv uses LOOCV by default
idw_model.cv <- gstat.cv(idw_model.model, nfold = nrow(sf_observations)) 

```

```{r calc_nrmse_idw_model}

idw_model.rsme <- calc_rmse(idw_model.cv$residual)
idw_model.nrsme <- calc_nrmse(idw_model.rsme, sf_observations$windspeed)

print(sprintf(
  "IDW [n=5, idp=1.5]: rmse=%f nrmse=%f", 
  idw_model.rsme, 
  idw_model.nrsme
))

```

### 3.4 Ordinary Kriging
#### 3.4.1 Fit variogram model 
```{r fit_ordinary_kriging_variogram, warning=FALSE}

# Create variogram
od_kriging.vgm <- variogram(windspeed ~ 1, data = sf_observations)

# Fit spherical variogram model to calculate sill and range
od_kriging.vgm.fit <- fit.variogram(od_kriging.vgm, vgm(348.342080, "Sph", 40943.119350))


```

#### 3.4.2 Train ordinary kriging model

```{r train_universal_kriging_models}

od_kriging.model <- gstat(
  formula = windspeed ~ 1,
  data = sf_observations,
  nmax = 5,
  model = od_kriging.vgm.fit
)

```


#### 3.4.3 Predict surfaces with ordinary kriging models

```{r pred_universal_kriging_surfaces}

od_kriging.model.preds <- as.data.frame(predict(od_kriging.model, sf_receptors))

```


#### 3.4.4 Show predictions in map

```{r show_ordinry_kriging_surfaces}

show_prediction_map(od_kriging.model.preds$var1.pred)

```

- Results show little to no difference in range of values, therefor it can be assumed that impact of seasonal or anual mean on ordinary kriging is minimal. 

#### 3.4.5 Validate results

```{r validate_simple_kriging_model, echo=FALSE}

od_kriging.cv <- gstat.cv(od_kriging.model, nfold = nrow(sf_observations))

```

```{r calc_nrmse_simple_kriging_model}

od_kriging.rsme <- calc_rmse(od_kriging.cv$residual)
od_kriging.nrsme <- calc_nrmse(od_kriging.rsme, sf_observations$windspeed)

print(sprintf(
  "Ordinary kriging [n=5, vgm=(348.342080, 'Sph', 40943.119350)]: rmse=%f nrmse=%f", 
  od_kriging.rsme, 
  od_kriging.nrsme
))

```

### 3.5 Universal kriging
#### 3.5.1 Fit variogram model for universal kriging

```{r universal_kriging_model_prepare}

sf_observations_u <- sf_observations %>%
  mutate( 
    x = st_coordinates(.)[,1],
    y = st_coordinates(.)[,2]
  )

sf_receptors_u <- sf_receptors %>%
  mutate( 
    x = st_coordinates(.)[,1],
    y = st_coordinates(.)[,2]
  )
```

```{r fit_universal_kriging_variogram, warning=FALSE}

# Create variogram
u_kriging.vgm <- variogram(windspeed ~ x + y, data = sf_observations_u)

# Fit spherical variogram model to calculate sill and range
u_kriging.vgm.fit <- fit.variogram(u_kriging.vgm, vgm(292.1622, "Sph", 40943.12))

```


#### 3.5.2 Train universal kriging models

```{r universal_kriging_model}

u_kriging.model_linear <- gstat(
  formula = windspeed ~ x + y,
  data = sf_observations_u,
  nmax = 5,
  model = u_kriging.vgm.fit
)

u_kriging.model_cubic <- gstat(
  formula = windspeed ~ abs(x + y + (x^2) + (x * y) + (y^2) + (x^3) + ((x^2) * y) + (x * (y^2)) + (y^3)),
  data = sf_observations_u,
  nmax = 5,
  model = u_kriging.vgm.fit
)

```

#### 3.5.3 Predict surface using universal kriging model

```{r preds_universal_kriging}

u_kriging.model_linear.preds <- as.data.frame(predict(u_kriging.model_linear, sf_receptors_u))
u_kriging.model_cubic.preds <- as.data.frame(predict(u_kriging.model_cubic, sf_receptors_u))

```

#### 3.5.4 Show predictions in map

```{r show_universal_kriging_surfaces}

show_prediction_map(u_kriging.model_linear.preds$var1.pred)
show_prediction_map(u_kriging.model_cubic.preds$var1.pred)

```

#### 3.5.5 Validate results

```{r validate_universal_kriging}

u_kriging.cv.linear <- gstat.cv(u_kriging.model_linear, nfold = nrow(sf_observations_u))
u_kriging.cv.cubic <- gstat.cv(u_kriging.model_cubic, nfold = nrow(sf_observations_u))

```

```{r calc_nrsme_universal_kriging}

u_kriging.rsme.linear <- calc_rmse(u_kriging.cv.linear$residual)
u_kriging.rsme.cubic <- calc_rmse(u_kriging.cv.cubic$residual)

u_kriging.nrsme.linear <- calc_nrmse(u_kriging.rsme.linear, sf_observations_u$windspeed)
u_kriging.nrsme.cubic <- calc_nrmse(u_kriging.rsme.cubic, sf_observations_u$windspeed)

print("Universal kriging [n=5, vgm=(292.1622, 'Sph', 40943.12)]")

print(sprintf(
  "[model='Linear']: rmse=%f nrmse=%f", 
  u_kriging.rsme.linear, u_kriging.nrsme.linear
))

print(sprintf(
  "[model='Cubic']: rmse=%f nrmse=%f", 
  u_kriging.rsme.cubic, u_kriging.nrsme.cubic
))


```






