---
title: "Thesis Validation Overview"
author: "Arsalan Anwari"
date: "2023-06-16"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
setwd("C:/Users/arsal/Workspaces/thesis_project_f2")

```



# 1. Data preparation

## 1.1 Load libraries
```{r load_libs, message=FALSE, warning=FALSE, results='hide'}

options("rgdal_show_exportToProj4_warnings"="none")

library(sf) # used to load maps and manipulate spatial features
library(tidyverse) # used for piping functionality of data manipulation functions
library(gstat) # used for interpolation and validation of idw and kriging models
library(caret) # used for cross validation of other models
library(seewave) # used to calculate surface roughness

library(ggplot2)
library(gridExtra) # to create nxn facets for ggplot

library(conflicted)
conflict_prefer("filter", "dplyr")
conflict_prefer("lag", "dplyr")

source("mqrbf-model-simple.R")


```

## 1.2 Load data
```{r load_data}

sf_oberservations_train <- readRDS("Data/knmi_windspeed_observations_training_set.rds")
sf_observations_anual <- readRDS("Data/knmi_windspeed_complete_set.rds")
sf_receptors <- readRDS("Data/knmi_windspeed_receptors_training_set.rds")

ls_training_results <- readRDS("Data/dynamic_training_routine_training_results_annual_avg.rds")
ls_training_settings <- readRDS("Data/dynamic_training_routine_training_settings_annual_avg.rds")



```

## 1.3 Prepare data of training set

```{r prepare_data}

sf_oberservations_train <- sf_oberservations_train %>%
    mutate(
      x = st_coordinates(.)[,1],
      y = st_coordinates(.)[,2]
    )

df_training_results.trend_surface <- ls_training_results$trend_surface
df_training_results.mqrbf <- ls_training_results$mqrbf
df_training_results.idw <- ls_training_results$idw
df_training_results.od_kriging <- ls_training_results$od_kriging
df_training_results.u_kriging <- ls_training_results$u_kriging

```

```{r show_data}

head(df_training_results.trend_surface)
head(df_training_results.mqrbf)
head(df_training_results.idw)
head(df_training_results.od_kriging)
head(df_training_results.u_kriging)

```

```{r prepare_model_opts}

model_opt.trend_surface <- c("model = linear", "model = hyperbolic")
model_opt.mqrbf <- sprintf("smoothing factor = %f", ls_training_settings$mqrbf_smoothing_factor)
model_opt.idw <- sprintf("neighbors = %d", ls_training_settings$neighbor_range)
model_opt.od_kriging <- sprintf("neighbors = %d", ls_training_settings$neighbor_range)
model_opt.u_kriging <- c(
  sprintf("neighbors = %d, model = %s", 5, "linear"),
  sprintf("neighbors = %d, model = %s", 5, "hyperbolic"),
  sprintf("neighbors = %d, model = %s", 6, "linear"),
  sprintf("neighbors = %d, model = %s", 6, "hyperbolic"),
  sprintf("neighbors = %d, model = %s", 7, "linear"),
  sprintf("neighbors = %d, model = %s", 7, "hyperbolic")
)

```

```{r modify_training_results}

df_training_results.trend_surface <- df_training_results.trend_surface %>%
  rename(model_opt_id = model_opt, old_rmse = old_rsme, new_rmse = new_rsme) %>%
  mutate(
    model_opt_id = as.integer(model_opt_id),
    model_opt_name = model_opt.trend_surface[model_opt_id],
    train_iteration = as.integer(train_iteration),
    old_rmse = as.double(old_rmse),
    new_rmse = as.double(new_rmse),
    loss_value = as.double(loss_value),
    case_weight = as.double(weights)
  ) %>%
  select( model_opt_id, model_opt_name, train_iteration, case_weight, old_rmse, new_rmse, loss_value )

df_training_results.mqrbf <- df_training_results.mqrbf %>%
  rename(model_opt_id = model_opt, old_rmse = old_rsme, new_rmse = new_rsme) %>%
  mutate(
    model_opt_id = as.integer(model_opt_id),
    model_opt_name = model_opt.mqrbf[model_opt_id],
    train_iteration = as.integer(train_iteration),
    old_rmse = as.double(old_rmse),
    new_rmse = as.double(new_rmse),
    loss_value = as.double(loss_value),
    alpha_weight = as.double(weights)
  ) %>%
  select( model_opt_id, model_opt_name, train_iteration, alpha_weight, old_rmse, new_rmse, loss_value )

df_training_results.idw <- df_training_results.idw %>%
  rename(model_opt_id = model_opt, old_rmse = old_rsme, new_rmse = new_rsme) %>%
  mutate(
    model_opt_id = as.integer(model_opt_id),
    model_opt_name = model_opt.idw[model_opt_id],
    train_iteration = as.integer(train_iteration),
    old_rmse = as.double(old_rmse),
    new_rmse = as.double(new_rmse),
    loss_value = as.double(loss_value),
    idp = as.double(weights)
  ) %>%
  select( model_opt_id, model_opt_name, train_iteration, idp, old_rmse, new_rmse, loss_value )

df_training_results.od_kriging <- df_training_results.od_kriging %>%
  rename(model_opt_id = model_opt, old_rmse = old_rsme, new_rmse = new_rsme) %>%
  mutate(
    model_opt_id = as.integer(model_opt_id),
    model_opt_name = model_opt.od_kriging[model_opt_id],
    train_iteration = as.integer(train_iteration),
    old_rmse = as.double(old_rmse),
    new_rmse = as.double(new_rmse),
    loss_value = as.double(loss_value),
    weights = strsplit(weights, ",")
  ) %>%
  unnest_wider(weights, names_sep = "_") %>%
  mutate( 
    psill = as.double(weights_2), 
    range = as.double(weights_3) 
  ) %>%
  select( model_opt_id, model_opt_name, train_iteration, psill, range, old_rmse, new_rmse, loss_value )

df_training_results.u_kriging <- df_training_results.u_kriging %>%
  rename(model_opt_id = model_opt, old_rmse = old_rsme, new_rmse = new_rsme) %>%
  mutate(
    model_opt_id = as.integer(model_opt_id),
    model_opt_name = model_opt.u_kriging[model_opt_id],
    train_iteration = as.integer(train_iteration),
    old_rmse = as.double(old_rmse),
    new_rmse = as.double(new_rmse),
    loss_value = as.double(loss_value),
    weights = strsplit(weights, ",")
  ) %>%
  unnest_wider(weights, names_sep = "_") %>%
  mutate( 
    psill = as.double(weights_2), 
    range = as.double(weights_3) 
  ) %>%
  select( model_opt_id, model_opt_name, train_iteration, psill, range, old_rmse, new_rmse, loss_value )

```


```{r show_modify_training_results}

head(df_training_results.trend_surface)
head(df_training_results.mqrbf)
head(df_training_results.idw)
head(df_training_results.od_kriging)
head(df_training_results.u_kriging)

```

```{r andskjasndk}

df_training_results.u_kriging

```



# 2. Analyse RMSE results of training set
## 2.1 Find weights with lowest RMSE per model variation for training set 

```{r find_lowest_rmse_per_variation}

find_lowest_rmse_per_model_opt <- function(df){
  df_grouped_rmse <- df %>%
    group_by(model_opt_id) %>%
    summarise(old_rmse = min(old_rmse))

  df_found <- df %>%
    filter(new_rmse %in% df_grouped_rmse$old_rmse) %>%
    rename(rmse = new_rmse) %>%
    select(!c(model_opt_id, train_iteration, loss_value, old_rmse))
  
  if (nrow(df_found) == 0){
    df_found <- df %>%
      filter(old_rmse %in% df_grouped_rmse$old_rmse) %>%
      rename(rmse = old_rmse) %>%
      select(!c(model_opt_id, train_iteration, loss_value, new_rmse))
  }
  
  return(df_found)
}

```


```{r get_lowest_rmse_per_variation}

df_lowest_rmse_per_opt.trend_surface <- find_lowest_rmse_per_model_opt(df_training_results.trend_surface)
df_lowest_rmse_per_opt.mqrbf <- find_lowest_rmse_per_model_opt(df_training_results.mqrbf)
df_lowest_rmse_per_opt.idw <- find_lowest_rmse_per_model_opt(df_training_results.idw)
df_lowest_rmse_per_opt.od_kriging <- find_lowest_rmse_per_model_opt(df_training_results.od_kriging)
df_lowest_rmse_per_opt.u_kriging <- find_lowest_rmse_per_model_opt(df_training_results.u_kriging)

```

```{r show_lowest_rmse_per_variation}

df_lowest_rmse_per_opt.trend_surface
df_lowest_rmse_per_opt.mqrbf
df_lowest_rmse_per_opt.idw
df_lowest_rmse_per_opt.od_kriging
df_lowest_rmse_per_opt.u_kriging

```

## 2.2 Save results of combination of weight and variation with lowest RMSE per model

```{r find_lowest_rmse_combo}

save.weights_with_lowest_rmse_per_opt <- list(
  trend_surface = df_lowest_rmse_per_opt.trend_surface,
  mqrbf = df_lowest_rmse_per_opt.mqrbf,
  idw = df_lowest_rmse_per_opt.idw,
  od_kriging = df_lowest_rmse_per_opt.od_kriging,
  u_kriging = df_lowest_rmse_per_opt.u_kriging
)

saveRDS(save.weights_with_lowest_rmse_per_opt, "Data/weights_with_lowest_rmse_per_opt.rds")

```

## 2.3 Show fluctuations of new weights and its effect on RMSE

- Measurements of the OK model where chosen as they showed the highest amount of fluctuation per model variation

```{r prepare_fluctuation_measurements_}

df_training_flucts <- df_training_results.od_kriging %>%
  select(model_opt_id, train_iteration, new_rmse, psill) %>%
  filter(model_opt_id == 2) %>%
  rename(rmse = new_rmse, weight = psill)

df_training_flucts

summary(df_training_flucts$rmse)

```

```{r plot_fluctuation_measurements_idw}

plot_train_results <- df_training_flucts %>% 
  ggplot( aes(x = train_iteration, y = rmse) ) +
  geom_line(position = "identity", linewidth = 0.5) +
  scale_fill_grey(start = 0.1, end = 0.9) +
  labs(title = "Training results (Ordinary kriging)", subtitle = "neighbors = 6") +
  theme_bw()

plot_weight_effect <- df_training_flucts %>% 
  ggplot( aes(x = train_iteration, y = weight) ) +
  geom_line(position = "identity", linewidth = 0.5) +
  scale_y_continuous(position = "right") +
  scale_fill_grey(start = 0.1, end = 0.9) +
  labs(title = "Used weights (Ordinary kriging)", subtitle = "neighbors = 6") +
  theme_bw()


grid.arrange(plot_train_results, plot_weight_effect, ncol = 2)


```




# 3. Prepare datasets and functions for different metrics
## 3.1 Prepare dataset
### 3.1.1 Seasonal set
```{r define_season_set_days}

season_set_days.spring <- as.Date(c(
  "2017-03-01", "2017-03-08", "2017-03-15", "2017-03-22",
  "2017-04-01", "2017-04-08", "2017-04-15", "2017-04-22",
  "2017-05-01", "2017-05-08", "2017-05-15", "2017-05-22"
))

season_set_days.summer <- as.Date(c(
  "2017-06-01", "2017-06-08", "2017-06-15", "2017-06-22",
  "2017-07-01", "2017-07-08", "2017-07-15", "2017-07-22",
  "2017-08-01", "2017-08-08", "2017-08-15", "2017-08-22"
))

season_set_days.autumn <- as.Date(c(
  "2017-09-01", "2017-09-08", "2017-09-15", "2017-09-22",
  "2017-10-01", "2017-10-08", "2017-10-15", "2017-10-22",
  "2017-11-01", "2017-11-08", "2017-11-15", "2017-11-22"
))

season_set_days.winter <- as.Date(c(
  "2017-12-01", "2017-12-08", "2017-12-15", "2017-12-22",
  "2017-01-01", "2017-01-08", "2017-01-15", "2017-01-22",
  "2017-02-01", "2017-02-08", "2017-02-15", "2017-02-22"
))

```

```{r create_seasonal_set}

sf_seasonal_set.spring <- sf_observations_anual %>%
  filter( date %in% season_set_days.spring )

sf_seasonal_set.summer <- sf_observations_anual %>%
  filter( date %in% season_set_days.summer )

sf_seasonal_set.autumn <- sf_observations_anual %>%
  filter( date %in% season_set_days.autumn )

sf_seasonal_set.winter <- sf_observations_anual %>%
  filter( date %in% season_set_days.winter )

```

### 3.1.2 Annual set

```{r define_annual_set_days}

gen_annuals_set_days <- function(){
  start_date <- as.Date("2017-01-01")
  out <- rep(NA, times = 48)

  i <- 1
  for(m in 0:11){
    for(d in 0:3){
      new_date <- start_date + months(m) + days(d)
      out[i] <- format(new_date, "%Y-%m-%d")
      i <- i + 1
    }
  }

  return(as.Date(out))
}

annual_set_days <- gen_annuals_set_days()

print(annual_set_days)

```

```{r create_annual_set}

sf_annual_set <- sf_observations_anual %>%
  filter( date %in% annual_set_days )

```

### 3.1.3 Random set

```{r define_random_set_days}

gen_random_set_days <- function(){
  set.seed(54353)
  
  start_date <- as.Date("2017-01-01")
  out <- rep(NA, times = 48)

  for(i in 1:48){
    rand_day <- as.integer(runif(1, min=6, max=24)) #make sure that no same day as annual set can be picked
    rand_month <- as.integer(runif(1, min=0, max=11))
    new_date <- start_date + months(rand_month) + days(rand_day)
    
    while(new_date %in% out){
      rand_day <- as.integer(runif(1, min=6, max=24))
      rand_month <- as.integer(runif(1, min=0, max=11))
      new_date <- start_date + months(rand_month) + days(rand_day)
    }
  
    out[i] <- format(new_date, "%Y-%m-%d")
  }

  return(as.Date(out))
}

random_set_days <- gen_random_set_days()
print(random_set_days)

```

```{r create_random_set}

sf_random_set <- sf_observations_anual %>%
 filter( date %in% random_set_days )

```

### 3.1.4 Clustered set

```{r create_sample_map}

sf_sample_map <- sf_observations_anual %>%
  filter( date == as.Date("2017-01-01") )

sf_sample_map

```

```{r plot_sample_map}

sf_sample_map %>%
  ggplot() +
    geom_sf(alpha = 0.0) +
    geom_sf_text(aes(label = station_code), size = 3) +
    theme_void()

```


As you can see from the plot, the following stations are in or near the province of Zeeland:

	- 312, 316, 324, 331, 340, 315, 310, 313, 308, 319, 330, 343
	
```{r define_clustered_set_days}

gen_clustered_set_days <- function(){
  set.seed(35433)
  
  start_date <- as.Date("2017-01-01")
  out <- rep(NA, times = 48)

  for(i in 1:48){
    rand_day <- as.integer(runif(1, min=0, max=28))
    rand_month <- as.integer(runif(1, min=0, max=11))
    new_date <- start_date + months(rand_month) + days(rand_day)
    
    while(new_date %in% out){
      rand_day <- as.integer(runif(1, min=6, max=24))
      rand_month <- as.integer(runif(1, min=0, max=11))
      new_date <- start_date + months(rand_month) + days(rand_day)
    }
    
    out[i] <- format(new_date, "%Y-%m-%d")
  }

  return(as.Date(out))
}

clustered_set_days <- gen_clustered_set_days()
print(clustered_set_days)

```
	
```{r create_clustered_set}

clustered_set_stations <- c(312, 316, 324, 331, 340, 315, 310, 313, 308, 319, 330, 343)

sf_clustered_set <- sf_observations_anual %>%
  filter( date %in% clustered_set_days & station_code %in% clustered_set_stations )

```
	
### 3.1.5 Sparse set

Based on the map shown in 3.1.4 the following stations where selected from the center and north of the Netherlands for the sparse set:

- Center: 260, 269, 275, 375, 215, 240
- North: 270, 280, 279, 273, 267, 242

The stations 260 and 270 can be seen as the center point where de distance between them and the other points is maximized.
	
```{r create_sparse_set}

sparse_set_stations <- c(260, 269, 275, 375, 215, 240, 270, 280, 279, 273, 267, 242)

sf_sparse_set <- sf_observations_anual %>%
  filter( date %in% clustered_set_days & station_code %in% sparse_set_stations )

```


### 3.1.6 Check if sets have correct sizes	
	
```{r compare_set_sizes}

station_count <- length(unique(sf_observations_anual$station_code))
all_season_set_size <- (
  length(unique(sf_seasonal_set.spring$date)) +
  length(unique(sf_seasonal_set.summer$date)) + 
  length(unique(sf_seasonal_set.autumn$date)) + 
  length(unique(sf_seasonal_set.winter$date))
)

cat(sprintf(
  "
  Anual set days = %d
  Seasonal set days = %d
    Seasonal days [spring] = %d
    Seasonal days [summer] = %d
    Seasonal days [autumn] = %d
    Seasonal days [winter] = %d
  Random set days = %d
  Clustered set days = %d
  Sparse set days = %d
  ",
  length(unique(sf_annual_set$date)), 
  all_season_set_size,
    length(unique(sf_seasonal_set.spring$date)),
    length(unique(sf_seasonal_set.summer$date)),
    length(unique(sf_seasonal_set.autumn$date)),
    length(unique(sf_seasonal_set.winter$date)),
  length(unique(sf_random_set$date)),
  length(unique(sf_clustered_set$date)),
  length(unique(sf_sparse_set$date))
))


```


## 3.2 Prepare functions
### 3.2.1 RMSE function
```{r define_rmse_function}

rmse <- function(residuals) {
  return(sqrt(sum((residuals)^2)/length(residuals)))
}

```

### 3.2.2 SLOO-CV functions to calculate rmse

```{r rmse_func_models}

get_rmse.trend_surface <- function(observations, formula){
  if(formula == "windspeed ~ x + y"){
    case_weight <- 1.151302 
  } else{
    case_weight <- 1.000508
  }
  
  model <- train(
    as.formula(formula),
    method = "lm",
    data = observations,
    weights = (1 / windspeed) + case_weight,
    trControl = trainControl(method = "LOOCV")
  )
  
  residuals <- resid(model)
  
  return( rmse(residuals) )
  
}

get_rmse.mqrbf <- function(observations, opts = FALSE){
  
  model <- mqrbf(
    data = observations, 
    alpha_seed = 1.166434,
    smoothing_factor = 90.770461
  )
  
  resisuals <- train.mqrbf(model)$residuals
  
  return( rmse(resisuals) ) 
  
}

get_rmse.idw <- function(observations, opts = FALSE){
  
  model <- gstat(
    formula = windspeed ~ 1,
    data = observations,
    nmax = 6,
    set = list(idp = 1.028791)
  )
  
  { sink(nullfile()); cv <- gstat.cv(model, nfold = nrow(observations), verbose=FALSE); sink(); }
  
  residuals <- cv$residual
  
  return( rmse(residuals) )
}

get_rmse.kriging <- function(observations, formula){
  if(formula == "windspeed ~ 1"){
    weight_psill <- 0.3641014
    weight_range <- 36.41014
    vgm_formula <- formula
  } else{ 
    vgm_formula <- "windspeed ~ x + y"
    if(formula == "windspeed ~ x + y"){
      weight_psill <- 0.4390736
      weight_range <- 43.90736
    } else{
      weight_psill <- 0.5031153
      weight_range <- 50.31153
    }
  }
  
  kriging.vgm <- variogram(as.formula(vgm_formula), data = observations)
  kriging.vgm.fit <- suppressWarnings(fit.variogram(kriging.vgm, vgm("Sph")))
  
  base_psill <-  kriging.vgm.fit$psill[1]
  base_range <- kriging.vgm.fit$range[2]
  new_sill <- base_psill + weight_psill
  new_range <- base_range + weight_range
  
  vgm_model <- suppressWarnings(fit.variogram(
    kriging.vgm, 
    vgm(
      psill = new_sill, 
      model = "Sph", 
      range = new_range
    )
  ))
  
  model <- suppressWarnings(gstat(
    formula = as.formula(formula),
    data = observations,
    nmax = 6,
    model = vgm_model,
  ))
    
  { sink(nullfile()); cv <- gstat.cv(model, nfold = nrow(observations), verbose=FALSE); sink(); }
  
  residuals <- cv$residual
  
  return( rmse(residuals) )
}

```

### 3.2.3 Prediction funtions to calculate receptor variables

```{r surface_function_models}

get_preds.trend_surface <- function(observations, formula){
  if(formula == "windspeed ~ x + y"){
    case_weight <- 1.151302 
  } else{
    case_weight <- 1.000508
  }
  
  model <- lm(
    as.formula(formula),
    data = observations,
    weights = (1 / windspeed) + case_weight
  )
  
  surface <- predict(model, sf_receptors)
  
  return( surface )
}

get_preds.mqrbf <-  function(observations, opts = FALSE){
  
  model <- mqrbf(
    data = observations, 
    alpha_seed = 1.166434,
    smoothing_factor = 90.770461
  )
  
  surface <- predict.mqrbf(model, sf_receptors)
  
  return( surface )
  
}

get_preds.idw <- function(observations, opts = FALSE){
  
  sink(nullfile()) # Hide dumb gstat output by forcing all output to a nullfile
  
  model <- gstat(
    formula = windspeed ~ 1,
    data = observations,
    nmax = 6,
    set = list(idp = 1.028791)
  )
  
  surface <- predict(model, sf_receptors)$var1.pred
  
  sink() # Restore output to console
  
  return(surface)
  
}

get_preds.kriging <- function(observations, formula){
  
  if(formula == "windspeed ~ 1"){
    weight_psill <- 0.3641014
    weight_range <- 36.41014
    vgm_formula <- formula
  } else{ 
    vgm_formula <- "windspeed ~ x + y"
    if(formula == "windspeed ~ x + y"){
      weight_psill <- 0.4390736
      weight_range <- 43.90736
    } else{
      weight_psill <- 0.5031153
      weight_range <- 50.31153
    }
  }
  
  sink(nullfile()) # Hide dumb gstat output by forcing all output to a nullfile
  
  kriging.vgm <- variogram(as.formula(vgm_formula), data = observations)
  kriging.vgm.fit <- suppressWarnings(fit.variogram(kriging.vgm, vgm("Sph")))
  
  base_psill <-  kriging.vgm.fit$psill[1]
  base_range <- kriging.vgm.fit$range[2]
  new_sill <- base_psill + weight_psill
  new_range <- base_range + weight_range
  
  vgm_model <- suppressWarnings(fit.variogram(
    kriging.vgm, 
    vgm(
      psill = new_sill, 
      model = "Sph", 
      range = new_range
    )
  ))
  
  model <- suppressWarnings(gstat(
    formula = as.formula(formula),
    data = observations,
    nmax = 6,
    model = vgm_model,
  ))
  
  surface <- predict(model, sf_receptors)$var1.pred
  
  sink() # Restore output to console
  
  return(surface)
  
}
  
  

```

### 3.2.4 Calculate mean RMSE for given model in given time period for any set
```{r calc_mean_rmse}

calc_mean_rmse <- function(model_rmse_func, model_observations, model_rmse_func_opts){ 
  model_rmse_results <- c()
  for(filter_date in unique(model_observations$date)){
    observations <- model_observations %>% filter(date == filter_date)
    model_rmse <- model_rmse_func(observations, model_rmse_func_opts)
    model_rmse_results <- append(model_rmse_results, model_rmse)
  }
  return(mean(model_rmse_results))
}

```

### 3.2.5 Calculate max/med/madSR for given model for all days in random set
```{r calc_surface_roughness}

calc_surface_roughness <- function(model_pred_func, model_observations, model_pred_func_opts){ 
  model_sr_results <- c()
  for(filter_date in unique(model_observations$date)){
    observations <- model_observations %>% filter(date == filter_date)
    model_pred_result <- model_pred_func(observations, model_pred_func_opts)
    model_pred_result <- model_pred_result / sd(model_observations$windspeed)
    model_sr_result <- roughness(model_pred_result, std = TRUE)
    model_sr_results <- append(model_sr_results, model_sr_result)
  }
  return(list(
    max = max(model_sr_results),
    med = median(model_sr_results),
    mad = mad(model_sr_results)
  ))
}

```

### 3.2.6 Calculate mean surface difference for given model for all days in random set

```{r calc_mean_surface_diff}

calc_mean_surface_diff <- function(model_pred_func, model_observations, model_pred_func_opts){ 
  model_pred_diff_results <- c()
  for(filter_date in unique(model_observations$date)){
    observations <- model_observations %>% filter(date == filter_date)
    model_pred_result <- model_pred_func(observations, model_pred_func_opts)
    model_pred_diff_result <- abs( max(model_pred_result) - min(model_pred_result) )
    model_pred_diff_results <- append(model_pred_diff_results, model_pred_diff_result)
  }
  return(mean(model_pred_diff_results))
}

```

### 3.2.7 Calculate mean surface MAD for given model for all days in random set
```{r calc_mean_surface_mad}

calc_mean_surface_mad <- function(model_pred_func, model_observations, model_pred_func_opts){ 
  model_pred_mad_results <- c()
  for(filter_date in unique(model_observations$date)){
    observations <- model_observations %>% filter(date == filter_date)
    model_pred_result <- model_pred_func(observations, model_pred_func_opts)
    model_pred_mad_result <- mad(model_pred_result)
    model_pred_mad_results <- append(model_pred_mad_results, model_pred_mad_result)
  }
  return(mean(model_pred_mad_results))
}

```


## 3.3 Prepare model data
### 3.3.3 Model Names 

```{r define_model_names}

model_names <- c(
  "Trend surface (1st order)",
  "Trend surface (3rd order)",
  "MQ-RBF",
  "IDW",
  "Ordinary kriging",
  "Universal kriging (1st order)",
  "Universal kriging (3rd order)"
)

```

### 3.3.4 Regresion polynomial formulas

```{r define_polynomials}

polynomial_formulas <- c(
  "windspeed ~ 1",
  "windspeed ~ x + y", 
  "windspeed ~ x + y + I(x^2) + (x * y) + I(y^2) + I(x^3) + (I(x^2) * y) + (x * I(y^2)) + I(y^3)",
  "windspeed ~ abs(x + y + (x^2) + (x * y) + (y^2) + (x^3) + ((x^2) * y) + (x * (y^2)) + (y^3))"
)

```


# 4. Stability metrics
## 4.1 Seasonal-annual variability 
```{r define_sav}

metric.stability.sav <- function(){
  
  get_model_score <- function(model_name, model_rmse_func, model_rmse_func_opts = FALSE){
    
    print(sprintf("Calculating score for model %s ....", model_name))
    mean_rmse_annual <- calc_mean_rmse(model_rmse_func, sf_annual_set, model_rmse_func_opts)
    
    print("Calculating rmse for each season...")
    score <- max(c(
        abs( mean_rmse_annual - calc_mean_rmse(model_rmse_func, sf_seasonal_set.spring, model_rmse_func_opts) ),
        abs( mean_rmse_annual - calc_mean_rmse(model_rmse_func, sf_seasonal_set.summer, model_rmse_func_opts) ), 
        abs( mean_rmse_annual - calc_mean_rmse(model_rmse_func, sf_seasonal_set.autumn, model_rmse_func_opts) ), 
        abs( mean_rmse_annual - calc_mean_rmse(model_rmse_func, sf_seasonal_set.winter, model_rmse_func_opts) ) 
    ))
    
    print(sprintf("Done calculating score!"))
    cat("\r\n")
    
    return(score)
  }
  
  scores <-  c(
    get_model_score(model_names[1], get_rmse.trend_surface, polynomial_formulas[2]), # 1st order trend surface
    get_model_score(model_names[2], get_rmse.trend_surface, polynomial_formulas[3]), # 3rd order trend_surface
    get_model_score(model_names[3], get_rmse.mqrbf),
    get_model_score(model_names[4], get_rmse.idw),
    get_model_score(model_names[5], get_rmse.kriging, polynomial_formulas[1]), # ordinary kriging
    get_model_score(model_names[6], get_rmse.kriging, polynomial_formulas[2]), # 1st order universal kriging
    get_model_score(model_names[7], get_rmse.kriging, polynomial_formulas[4]) # 3rd order universal kriging
  )
  
  return( data.frame(model = model_names, score = scores) )
    
}
  
```

```{r calc_sav}

results.metric.stability.sav = metric.stability.sav()

```

```{r show_sav}

results.metric.stability.sav %>% arrange(score)

```

## 4.2 Seasonal-seasonal variability

```{r define_ssv}

metric.stability.ssv <- function(){
  
  get_model_score <- function(model_name, model_rmse_func, model_rmse_func_opts = FALSE){
    
    print(sprintf("Calculating score for model %s ....", model_name))
    
    print("Calculating rmse for each season...")
    scores <- c(
        calc_mean_rmse(model_rmse_func, sf_seasonal_set.spring, model_rmse_func_opts),
        calc_mean_rmse(model_rmse_func, sf_seasonal_set.summer, model_rmse_func_opts), 
        calc_mean_rmse(model_rmse_func, sf_seasonal_set.autumn, model_rmse_func_opts), 
        calc_mean_rmse(model_rmse_func, sf_seasonal_set.winter, model_rmse_func_opts)
    )
    
    score <- abs( max(scores) - min(scores) )
    
    print(sprintf("Done calculating score!"))
    cat("\r\n")
    
    return(score)
  }
  
  scores <-  c(
    get_model_score(model_names[1], get_rmse.trend_surface, polynomial_formulas[2]), # 1st order trend surface
    get_model_score(model_names[2], get_rmse.trend_surface, polynomial_formulas[3]), # 3rd order trend_surface
    get_model_score(model_names[3], get_rmse.mqrbf),
    get_model_score(model_names[4], get_rmse.idw),
    get_model_score(model_names[5], get_rmse.kriging, polynomial_formulas[1]), # ordinary kriging
    get_model_score(model_names[6], get_rmse.kriging, polynomial_formulas[2]), # 1st order universal kriging
    get_model_score(model_names[7], get_rmse.kriging, polynomial_formulas[4]) # 3rd order universal kriging
  )
  
  return( data.frame(model = model_names, score = scores) )
    
}
  
```

```{r calc_ssv}

results.metric.stability.ssv <- metric.stability.ssv()

```

```{r show_ssv}

results.metric.stability.ssv %>% arrange(score)

```

## 4.3 Annual random variability 

```{r define_arv}

metric.stability.arv <- function(){
  
  get_model_score <- function(model_name, model_rmse_func, model_rmse_func_opts = FALSE){
    
    print(sprintf("Calculating score for model %s ....", model_name))
    
    mean_rmse_annual <- calc_mean_rmse(model_rmse_func, sf_annual_set, model_rmse_func_opts)
    mean_rmse_random <- calc_mean_rmse(model_rmse_func, sf_random_set, model_rmse_func_opts)
    
    score <- abs(mean_rmse_annual - mean_rmse_random)
    
    print(sprintf("Done calculating score!"))
    cat("\r\n")
    
    return(score)
  }
  
  scores <-  c(
    get_model_score(model_names[1], get_rmse.trend_surface, polynomial_formulas[2]), # 1st order trend surface
    get_model_score(model_names[2], get_rmse.trend_surface, polynomial_formulas[3]), # 3rd order trend_surface
    get_model_score(model_names[3], get_rmse.mqrbf),
    get_model_score(model_names[4], get_rmse.idw),
    get_model_score(model_names[5], get_rmse.kriging, polynomial_formulas[1]), # ordinary kriging
    get_model_score(model_names[6], get_rmse.kriging, polynomial_formulas[2]), # 1st order universal kriging
    get_model_score(model_names[7], get_rmse.kriging, polynomial_formulas[4]) # 3rd order universal kriging
  )
  
  return( data.frame(model = model_names, score = scores) )
    
}
  
```

```{r calc_arv}

results.metric.stability.arv <- metric.stability.arv()

```

```{r show_arv}

results.metric.stability.arv %>% arrange(score)

```

## 4.4 Random-random variability 

```{r define_rrv}

metric.stability.rrv <- function(){
  
  get_model_score <- function(model_name, model_rmse_func, model_rmse_func_opts = FALSE){
    
    print(sprintf("Calculating score for model %s ....", model_name))
    
    dates <- unique(sf_random_set$date)
    date_chunks <- split(dates, rep(1:4, length.out = length(dates)))
    names(date_chunks) <- paste0("chunk_", 1:4)
    
    random_set.chunk_1 <- sf_random_set %>% filter(date %in% date_chunks$chunk_1)
    random_set.chunk_2 <- sf_random_set %>% filter(date %in% date_chunks$chunk_2)
    random_set.chunk_3 <- sf_random_set %>% filter(date %in% date_chunks$chunk_3)
    random_set.chunk_4 <- sf_random_set %>% filter(date %in% date_chunks$chunk_4)
    
    scores <- c(
        calc_mean_rmse(model_rmse_func, random_set.chunk_1, model_rmse_func_opts),
        calc_mean_rmse(model_rmse_func, random_set.chunk_2, model_rmse_func_opts), 
        calc_mean_rmse(model_rmse_func, random_set.chunk_3, model_rmse_func_opts), 
        calc_mean_rmse(model_rmse_func, random_set.chunk_4, model_rmse_func_opts)
    )
    
    score <- abs( max(scores) - min(scores) )
    
    print(sprintf("Done calculating score!"))
    cat("\r\n")
    
    return(score)
  }
  
  scores <-  c(
    get_model_score(model_names[1], get_rmse.trend_surface, polynomial_formulas[2]), # 1st order trend surface
    get_model_score(model_names[2], get_rmse.trend_surface, polynomial_formulas[3]), # 3rd order trend_surface
    get_model_score(model_names[3], get_rmse.mqrbf),
    get_model_score(model_names[4], get_rmse.idw),
    get_model_score(model_names[5], get_rmse.kriging, polynomial_formulas[1]), # ordinary kriging
    get_model_score(model_names[6], get_rmse.kriging, polynomial_formulas[2]), # 1st order universal kriging
    get_model_score(model_names[7], get_rmse.kriging, polynomial_formulas[4]) # 3rd order universal kriging
  )
  
  return( data.frame(model = model_names, score = scores) )
    
}
  
```

```{r calc_rrv}

results.metric.stability.rrv <- metric.stability.rrv()

```

```{r show_rrv}

results.metric.stability.rrv %>% arrange(score)

```

## 4.5 Global-local variability 

```{r define_glv}

metric.stability.glv <- function(){
  
  get_model_score <- function(model_name, model_rmse_func, model_rmse_func_opts = FALSE){
    
    print(sprintf("Calculating score for model %s ....", model_name))
    
    mean_rmse_clustered <- calc_mean_rmse(model_rmse_func, sf_clustered_set, model_rmse_func_opts)
    mean_rmse_sparse <- calc_mean_rmse(model_rmse_func, sf_sparse_set, model_rmse_func_opts)
    
    score <- abs(mean_rmse_clustered - mean_rmse_sparse)
    
    print(sprintf("Done calculating score!"))
    cat("\r\n")
    
    return(score)
  }
  
  scores <-  c(
    get_model_score(model_names[1], get_rmse.trend_surface, polynomial_formulas[2]), # 1st order trend surface
    get_model_score(model_names[2], get_rmse.trend_surface, polynomial_formulas[3]), # 3rd order trend_surface
    get_model_score(model_names[3], get_rmse.mqrbf),
    get_model_score(model_names[4], get_rmse.idw),
    get_model_score(model_names[5], get_rmse.kriging, polynomial_formulas[1]), # ordinary kriging
    get_model_score(model_names[6], get_rmse.kriging, polynomial_formulas[2]), # 1st order universal kriging
    get_model_score(model_names[7], get_rmse.kriging, polynomial_formulas[4]) # 3rd order universal kriging
  )
  
  return( data.frame(model = model_names, score = scores) )
    
}

```

```{r calc_glv}

results.metric.stability.glv <- metric.stability.glv()

```
```{r show_glv}

results.metric.stability.glv %>% arrange(score)

```

## 4.6 Compare results of all metrics

```{r compare_stability_metric_scores}

results.metric.stability.all <- results.metric.stability.sav
colnames(results.metric.stability.all) <- c("model", "sav_score")

results.metric.stability.all$ssv_score <- results.metric.stability.ssv$score
results.metric.stability.all$arv_score <- results.metric.stability.arv$score
results.metric.stability.all$rrv_score <- results.metric.stability.rrv$score
results.metric.stability.all$glv_score <- results.metric.stability.glv$score

results.metric.stability.all

```




# 5. Surface metrics

## 5.1 Surface roughness (MaxSR, MedSR and MadSR)

```{r define_max_med_mad_sr}

metric.surface.sr <- function(){

  max_sr_results <- c()
  med_sr_results <- c()
  mad_sr_results <- c()
  
  calc_model_score <- function(model_name, model_pred_func, model_pred_func_opts = FALSE){
    print(sprintf("Calculating surface roughness for model %s ....", model_name))
    results <- calc_surface_roughness(model_pred_func, sf_observations_anual, model_pred_func_opts)
    
    max_sr_results <<- append(max_sr_results, results$max)
    med_sr_results <<- append(med_sr_results, results$med)
    mad_sr_results <<- append(mad_sr_results, results$mad)
    
    print(sprintf("Done calculating surface roughness!"))
    cat("\r\n")
  }
  
  calc_model_score(model_names[1], get_preds.trend_surface, polynomial_formulas[2]) # 1st order trend surface
  calc_model_score(model_names[2], get_preds.trend_surface, polynomial_formulas[3]) # 3rd order trend_surface
  calc_model_score(model_names[3], get_preds.mqrbf)
  calc_model_score(model_names[4], get_preds.idw)
  calc_model_score(model_names[5], get_preds.kriging, polynomial_formulas[1]) # ordinary kriging
  calc_model_score(model_names[6], get_preds.kriging, polynomial_formulas[2]) # 1st order universal kriging
  calc_model_score(model_names[7], get_preds.kriging, polynomial_formulas[4]) # 3rd order universal kriging

  return( data.frame(
    model = model_names,
    max_sr = max_sr_results,
    med_sr = med_sr_results,
    mad_sr = mad_sr_results
  ))

}


```
 
```{r calc_max_med_mad_sr}

results.metric.surface.sr <- metric.surface.sr()

```
 
```{r show_max_med_mad_sr}

results.metric.surface.sr %>% arrange(mad_sr)

```

## 5.2 Minimum maximum surface deviation

```{r define_mi_ma_sd}

metric.surface.mi_ma_sd <- function(){

  get_model_score <- function(model_name, model_pred_func, model_pred_func_opts = FALSE){
    print(sprintf("Calculating MiMaSD score for model %s ....", model_name))
    
    score <- calc_mean_surface_diff(model_pred_func, sf_observations_anual, model_pred_func_opts)
    
    print(sprintf("Done calculating MiMaSD score!"))
    cat("\r\n")
    
    return(score)
  }
  
  scores <- c(
    get_model_score(model_names[1], get_preds.trend_surface, polynomial_formulas[2]), # 1st order trend surface
    get_model_score(model_names[2], get_preds.trend_surface, polynomial_formulas[3]), # 3rd order trend_surface
    get_model_score(model_names[3], get_preds.mqrbf),
    get_model_score(model_names[4], get_preds.idw),
    get_model_score(model_names[5], get_preds.kriging, polynomial_formulas[1]), # ordinary kriging
    get_model_score(model_names[6], get_preds.kriging, polynomial_formulas[2]), # 1st order universal kriging
    get_model_score(model_names[7], get_preds.kriging, polynomial_formulas[4]) # 3rd order universal kriging
  )
  return( data.frame(model = model_names, score = scores) )

}

```
 
```{r calc_mi_ma_sd}

results.metric.surface.mi_ma_sd <- metric.surface.mi_ma_sd()

```

```{r show_mi_ma_sd}

results.metric.surface.mi_ma_sd %>% arrange(score)

```
 
 ## 5.3 MAD surface deviation
 
```{r define_mad_sd}

metric.surface.mad_sd <- function(){

  get_model_score <- function(model_name, model_pred_func, model_pred_func_opts = FALSE){
    print(sprintf("Calculating MadSD score for model %s ....", model_name))
    
    score <- calc_mean_surface_mad(model_pred_func, sf_observations_anual, model_pred_func_opts)
    
    print(sprintf("Done calculating MadSD score!"))
    cat("\r\n")
    
    return(score)
  }
  
  scores <- c(
    get_model_score(model_names[1], get_preds.trend_surface, polynomial_formulas[2]), # 1st order trend surface
    get_model_score(model_names[2], get_preds.trend_surface, polynomial_formulas[3]), # 3rd order trend_surface
    get_model_score(model_names[3], get_preds.mqrbf),
    get_model_score(model_names[4], get_preds.idw),
    get_model_score(model_names[5], get_preds.kriging, polynomial_formulas[1]), # ordinary kriging
    get_model_score(model_names[6], get_preds.kriging, polynomial_formulas[2]), # 1st order universal kriging
    get_model_score(model_names[7], get_preds.kriging, polynomial_formulas[4]) # 3rd order universal kriging
  )
  return( data.frame(model = model_names, score = scores) )

}

```
 

```{r calc_mid_sd}

results.metric.surface.mad_sd <- metric.surface.mad_sd()

```

```{r show_mad_sd}

results.metric.surface.mad_sd %>% arrange(score)

```
 
## 5.4 Compare results of all metrics
```{r compare_surface_metric_scores}

results.metric.surface.all <- results.metric.surface.sr

results.metric.surface.all$mi_mad_sd <- results.metric.surface.mi_ma_sd$score
results.metric.surface.all$mad_sd <- results.metric.surface.mad_sd$score

results.metric.surface.all

```